---
title: "DATA607 Assignment 3"
author: "Albert Gilharry"
date: "15 February 2018"
output:
  pdf_document:
    toc: yes
  html_document:
    css: ./css.css
    highlight: pygments
    pdf_document: default
    theme: spacelab
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br />
<br />

```{r}
library("stringr")
```
<br />

3. Copy the introductory example. The vector *name* stores the extracted names.


```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555 -6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```
<br />

  - Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.
  
      
  <div id="solution">
      
       Assuming that this question refers to names that are in reversed order and separated by a comma, and that titles and other designations should remain in place:
        
    ```{r}
    names <- str_split(name,", ")
    
    for( i in 1:length(name) ){
       if(length(names[[i]]) == 2){
         name[i]<-str_c(names[[i]][2], names[[i]][1], sep=" ")
       }
    }
    
    ```
      
    Result:  
    ```{r}
    print(name)
    ```
      
  </div>
  <br />
      
  - Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
  
    
  <div id="solution">
    
    My assumption here is that titles start with a single uppercase letter followed by 1 to 3 alphabetic characters and concludes with a dot.
    
    ```{r}
    str_detect(name,"[:upper:]{1}[:alpha:]{1,3}\\.")
    ```
    
  </div>
  <br />
      
  - Construct a logical vector indicating whether a character has a second name.
  
    
  <div id="solution">
    
    My approach to this is to separate the expression into two categories, people with a title and second name and people without a title but has a second name. I assumed that a title must start with an uppercase character and must be followed 1 to 3 aplhabetic characters and ends with a dot. An uppercase letter followed by dot will be treated as an abbreviated first name. These conditions are separated by a pipe to accomodate with both types. A person with a title and a second name will has an additional space followed by alphabetic characters in their string.
      
    ```{r}
      str_detect(name,"([:upper:]{1}\\.([:space:][:upper:][:alpha:]+){2})|([:upper:]{1}[:lower:]+\\.([:space:][:upper:][:alpha:]+){3})")
    ```
    
  </div>
  <br />
    
4.  Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

  - [0-9]+\\$
  
      
  <div id="solution">
      
        This expression represents numbers with a minimum length of 1 digit followed by a dollar sign *$*.
        
        Example: *0235$*
        
    ```{r}
    str_extract("0235$","[0-9]+\\$")
    ```
  </div>
  <br />
      
  - \\b[a-z]{1,4}\\b
  
  
  <div id="solution">
      
      This expression represents words that are up to four letters long.
      
      Example: *a word*
      
    ```{r}
    str_extract_all("a word","\\b[a-z]{1,4}\\b")
    ```
  </div>
  <br />
  
  - .*?\\.txt$
  
  
  <div id="solution">
      
      This expression represents strings starts with 0 or more characters and ends with *.txt*. This is representative of files with the extension *.txt*, but having an actual name for the file is optional.
      
      Example: *.txt*
      
    ```{r}
    str_detect(".txt",".*?\\.txt$")
    ```
    
  </div>
  <br />
  
      
  - \\d{2}/\\d{2}/\\d{4}
  
  <div id="solution">
      This represents 8 digits separated by a forward slash after the second digit and one after the fouth digit.
      
      Example: 24/65/9987
      
    ```{r}
    str_extract("29/22/2333","\\d{2}/\\d{2}/\\d{4}")
    ```
    
  </div>
  <br />
  
  - <(.+?)>.+?</\\1>
  
  <div id="solution">
        The represents a text with at minimum 1 character enclosed in an XML type tag. The opening must be closed following the enclosed text.
        Example: <d>x</d>
    ```{r}
    str_extract("<d>x</d>","<(.+?)>.+?</\\1>")
    ``` 
        
  </div>
  <br />
      
9.  The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

<div>
clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3c
oc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.
SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr
</div>

   <div id="solution">
     
     Extract the uppercase letters and punctuations from the text.
     
```{r}
txt <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

str_extract_all(txt,"[:upper:]|[:punct:]")

  
```
    
    This reveals "CONGRATULATIONS YOU ARE A SUPERNERD!".
  </div>